Manuscript draft

### 2019-08-23 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@

Begin discussion by emphasising that this these conclusions are based on analyses of an unprecedented amount of data across many fluxes, limited to mature undisturbed forests.

>>>>

In this analysis we use a comprehensive global database, standardised to account for differences in methodology across studies, and stand age and level of disturbance across sites. The database provides us with an unprecedented amount of data, representing all (?) global forest ecosystems and all significant forest carbon fluxes. This breadth of data allows us to comprehensively explore the relationships between climate and productivity on a global scale.

Past studies have differed in their conclusions regarding the relationship between productivity and latitude. Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.

There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects. We find that temperature-related climate variables explain the highest proportion of variability in productivity.

There is debate over whether productivity in regions with higher MAT is higher because productivity inherently increases with air temperature during growing season months, or because higher MAT is correlated with longer growing seasons. Our analysis suggests that length of growing season is a key factor in understanding latitudinal patterns in productivity. We find that length of growing season is strongly correlated with productivity. This is supported by the strong correlation between temperature seasonality and productivity, indicating that aseasonality and year-round growing seasons have a strong influence over productivity.

These conclusions are supported by our analyses of the influence of climate within the growing season. We find much weaker correlations between climate variables and productivity when only growing season months are considered. There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes. This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT. However, we note that length of growing season is a less good predictor of productivity than MAT. This may be an artefact of imprecise estimation of length of growing season, and with more precise climate data on finer time-scales it is likely that the predictive ability of length of growing season would increase, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain. It is hard to attribute mechanisms through such coarse-scale data as this. Future studies.......

The majority of studies to date have focused on MAT and MAP as two key climate variables for which data is easily obtained. We find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature. This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature. This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to understand patterns of productivity. This is likely to be significant in the future, as the effects of a warming climate will be mediated through changes in precipitation regime.

This studies considers the relationship between productivity and climate on a coarse scale. Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites. As a result the climate data we used was primarily multi-year averages of annual climate measures from global databases. While the quality of these data is very high, we cannot expect that it captures detailed local variation in climate. Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT. With high-quality stand-level data for other climate variables we might find stronger predictors. Similarly, with accurate stand-level measurements of growing season length, we would expect to record a stronger correlation between growing season length and productivity.

Furthermore, we expect that forest stand productivity varies on a seasonal and annual scale with seasonal and annual variation in climate. By taking long-term averages both in climate and in forest productivity estimates, we expect that much of this variation will be averaged out, and the strong local effects of climate on productivity will be masked.

By testing a wide selection of carbon fluxes, we have been able to explore the relationships between fluxes and their responses to productivity. We see that:
- All fluxes show similar responses to climate drivers
- Allocation to fluxes does not vary with climate on a global scale. This suggests that the pressures that require individuals to shift allocation between fluxes occur on much more local scales. We expect that allocation varies with factors such as altitude, stand age, disturbance regime, micro-climate, and is likely to show seasonal responses to climate variables, such as showing short-term responses to water availability.
- Climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes. This indicates that factors other than climate influence subsidiary fluxes. This links to the results we find for allocation: although climate has a significant effect on these fluxes, there are a number of other factors which can influence the way that carbon is allocated to them, which masks the effect of climate to some extent.

@@
